---
title: "Week 8 Ex Solution"
author: "Katz"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
```


For this week's exercise, we will be looking at differences in students' SAT scores.

We will simulate data that we might have where a dependent samples t-test would be appropriate. 

We are going to simulate the scenario where we have student SAT scores on both the verbal section and math section.

```{r}
samp_size <- 1000

student_id <- seq(samp_size)
sat_verbal <- round(rnorm(n = samp_size, mean = 620, sd = 55), 0)
sat_math <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

sat_verbal_capped <- modify(.x = sat_verbal, .f = ~ min(800, .x))
sat_math_capped <- modify(.x = sat_math, .f = ~min(800, .x))

sat_df <- tibble(id = student_id,
                 verbal = sat_verbal_capped,
                 math = sat_math_capped)

```

Up to now, we only simulated data and stored it in a dataframe called sat_df. 

For your first step, try visualizing the verbal and math scores using histograms.

1. Verbal scores
```{r}

sat_df %>% ggplot(aes(x = verbal)) +
  geom_histogram() +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")
```

2. Math scores
```{r}


sat_df %>% ggplot(aes(x = math)) +
  geom_histogram() +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```


Alternatively, try to plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter).

```{r}
sat_df %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 

Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test.

```{r}
sat_t_test <- t.test(sat_df$verbal, sat_df$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test

```


We see there was a statistically significant difference, but now let's try changing the distributions and see what happens.

First, we'll change make the means very close together and see how sensitive the test is to that

```{r}
samp_size <- 1000

student_id_2 <- seq(samp_size)
sat_verbal_2 <- round(rnorm(n = samp_size, mean = 690, sd = 55), 0)
sat_math_2 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

# let's cap the simulated scores at 800
sat_verbal_2_capped <- modify(.x = sat_verbal_2, .f = ~ min(800, .x))
sat_math_2_capped <- modify(.x = sat_math_2, .f = ~min(800, .x))


sat_df_2 <- tibble(id = student_id,
                 verbal = sat_verbal_2_capped,
                 math = sat_math_2_capped)

sat_df_2 %>% ggplot(aes(x = verbal)) +
  geom_histogram() +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")
```

```{r}

sat_df_2 %>% ggplot(aes(x = math)) +
  geom_histogram() +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```


Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter).

```{r}
sat_df_2 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```


Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 

Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test
```{r}
sat_t_test_2 <- t.test(sat_df_2$verbal, sat_df_2$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_2

```




Let's try a third test - this time we'll change the sample size to be 25 students rather than 1000

```{r}
samp_size <- 25

student_id_3 <- seq(samp_size)
sat_verbal_3 <- round(rnorm(n = samp_size, mean = 690, sd = 55), 0)
sat_math_3 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

# cap the scores at 800
sat_verbal_3_capped <- modify(.x = sat_verbal_3, .f = ~ min(800, .x))
sat_math_3_capped <- modify(.x = sat_math_3, .f = ~min(800, .x))



sat_df_3 <- tibble(id = student_id_3,
                 verbal = sat_verbal_3_capped,
                 math = sat_math_3_capped)

sat_df_3 %>% ggplot(aes(x = verbal)) +
  geom_histogram() +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")


sat_df_3 %>% ggplot(aes(x = math)) +
  geom_histogram() +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")


```

Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter)

```{r}
sat_df_3 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct.Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test

```{r}
sat_t_test_3 <- t.test(sat_df_3$verbal, sat_df_3$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_3

```


Now we see that the statistically significant difference is gone. How about if we go back to having a large difference in the means

```{r}
samp_size <- 25

student_id_4 <- seq(samp_size)
sat_verbal_4 <- round(rnorm(n = samp_size, mean = 620, sd = 55), 0)
sat_math_4 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

#cap the scores at 800
sat_verbal_4_capped <- modify(.x = sat_verbal_4, .f = ~ min(800, .x))
sat_math_4_capped <- modify(.x = sat_math_4, .f = ~min(800, .x))


sat_df_4 <- tibble(id = student_id_4,
                   verbal = sat_verbal_4_capped,
                   math = sat_math_4_capped)

# visualize the distributions of scores
sat_df_4 %>% ggplot(aes(x = verbal)) +
  geom_histogram() +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")


sat_df_4 %>% ggplot(aes(x = math)) +
  geom_histogram() +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```

Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter)

```{r}
sat_df_4 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 
Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test

```{r}
sat_t_test_4 <- t.test(sat_df_4$math, sat_df_4$verbal, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_4
```


We should see that the t-test will pick up differences between means if there is a sufficiently large difference.