---
title: "Week 8 Ex Solution"
author: "Katz"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
```


For this week's exercise, we will be looking at differences in students' SAT scores.

We will simulate data that we might have where a dependent samples t-test would be appropriate. 

We are going to simulate the scenario where we have student SAT scores on both the verbal section and math section.

```{r}
samp_size <- 1000

student_id <- seq(samp_size)
sat_verbal <- round(rnorm(n = samp_size, mean = 620, sd = 55), 0)
sat_math <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

sat_verbal_capped <- modify(.x = sat_verbal, .f = ~ min(800, .x))
sat_math_capped <- modify(.x = sat_math, .f = ~min(800, .x))

sat_df <- tibble(id = student_id,
                 verbal = sat_verbal_capped,
                 math = sat_math_capped)

```

Up to now, we only simulated data and stored it in a dataframe called sat_df. 

For your first step, try visualizing the verbal and math scores using histograms.

1. Verbal scores
```{r}

sat_df %>% ggplot(aes(x = verbal)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")

```

2. Math scores
```{r}


sat_df %>% ggplot(aes(x = math)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```


Alternatively, try to plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter).

```{r}
sat_df %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), color = "white", alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), color = "white", alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

You should also try doing this with a boxplot. I give you the code for this first version since it involves a `pivot_longer()`, but you should see if you can replicate it for the next rounds on your own.

```{r}
sat_df %>% 
  pivot_longer(cols = c("verbal", "math"), 
               names_to = "variable",
               values_to = "score") %>% 
  ggplot(aes(x = score, y = variable)) +
  geom_boxplot() +
  labs(x = "SAT Test Section",
       y = "Section Score",
       title = "Distribution of SAT Math and Verbal Scores")


```



Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 

Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test.

```{r}
sat_t_test <- t.test(sat_df$verbal, sat_df$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test

```

```{r}

(t <- sat_t_test$statistic[[1]]) # wrapping this in parentheses prints out the value of t at the same time that it stores the value of t

(df <- sat_t_test$parameter[[1]])

(r <- sqrt(t^2/(t^2 + df)))
round(r,3)
r^2
```





We see there was a statistically significant difference, but now let's try changing the distributions and see what happens.


### Round 2


First, we'll change make the means very close together and see how sensitive the test is to that

```{r}
samp_size <- 1000

student_id_2 <- seq(samp_size)
sat_verbal_2 <- round(rnorm(n = samp_size, mean = 690, sd = 55), 0)
sat_math_2 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

# let's cap the simulated scores at 800
sat_verbal_2_capped <- modify(.x = sat_verbal_2, .f = ~ min(800, .x))
sat_math_2_capped <- modify(.x = sat_math_2, .f = ~min(800, .x))


sat_df_2 <- tibble(id = student_id,
                 verbal = sat_verbal_2_capped,
                 math = sat_math_2_capped)
```



```{r}
sat_df_2 %>% ggplot(aes(x = verbal)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")
```

```{r}

sat_df_2 %>% ggplot(aes(x = math)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```


Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter).

```{r}
sat_df_2 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), color = "white", alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), color = "white", alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

You should also try doing this with a boxplot.

```{r}
sat_df_2 %>% 
  pivot_longer(cols = c("verbal", "math"), 
               names_to = "variable",
               values_to = "score") %>% 
  ggplot(aes(x = variable, y = score)) +
  geom_boxplot() +
  labs(x = "SAT Test Section",
       y = "Section Score",
       title = "Distribution of SAT Math and Verbal Scores")

```


Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 

Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test
```{r}
sat_t_test_2 <- t.test(sat_df_2$verbal, sat_df_2$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_2

```


Calculate an effect size. Remember that $r = \sqrt{\frac{t^2}{t^2 + df}}$

```{r}

(t <- sat_t_test_2$statistic[[1]]) # wrapping this in parentheses prints out the value of t at the same time that it stores the value of t

(df <- sat_t_test_2$parameter[[1]])

(r <- sqrt(t^2/(t^2 + df)))
round(r,3)
r^2
```




### Round 3


Let's try a third test - this time we'll change the sample size to be 25 students rather than 1000

```{r}
samp_size <- 25

student_id_3 <- seq(samp_size)
sat_verbal_3 <- round(rnorm(n = samp_size, mean = 690, sd = 55), 0)
sat_math_3 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

# cap the scores at 800
sat_verbal_3_capped <- modify(.x = sat_verbal_3, .f = ~ min(800, .x))
sat_math_3_capped <- modify(.x = sat_math_3, .f = ~min(800, .x))



sat_df_3 <- tibble(id = student_id_3,
                 verbal = sat_verbal_3_capped,
                 math = sat_math_3_capped)
```


```{r}
sat_df_3 %>% ggplot(aes(x = verbal)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")
```


```{r}
sat_df_3 %>% ggplot(aes(x = math)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")


```

Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter)

```{r}
sat_df_3 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), color = "white", alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), color = "white", alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

Try making a boxplot instead of histogram.

```{r}
sat_df_3 %>% 
  pivot_longer(cols = c("verbal", "math"), 
               names_to = "variable",
               values_to = "score") %>% 
  ggplot(aes(x = variable, y = score)) +
  geom_boxplot() +
  labs(x = "SAT Test Section",
       y = "Section Score",
       title = "Distribution of SAT Math and Verbal Scores")

```



Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct.Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test

```{r}
sat_t_test_3 <- t.test(sat_df_3$verbal, sat_df_3$math, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_3

```



Calculate an effect size. Remember that $r = \sqrt{\frac{t^2}{t^2 + df}}$

```{r}

(t <- sat_t_test_3$statistic[[1]]) # wrapping this in parentheses prints out the value of t at the same time that it stores the value of t

(df <- sat_t_test_3$parameter[[1]])

(r <- sqrt(t^2/(t^2 + df)))
round(r,3)
r^2
```











### Round 4


Now we see that the statistically significant difference is gone. How about if we go back to having a large difference in the means

```{r}
samp_size <- 25

student_id_4 <- seq(samp_size)
sat_verbal_4 <- round(rnorm(n = samp_size, mean = 620, sd = 55), 0)
sat_math_4 <- round(rnorm(n = samp_size, mean = 700, sd = 40), 0)

#cap the scores at 800
sat_verbal_4_capped <- modify(.x = sat_verbal_4, .f = ~ min(800, .x))
sat_math_4_capped <- modify(.x = sat_math_4, .f = ~min(800, .x))


sat_df_4 <- tibble(id = student_id_4,
                   verbal = sat_verbal_4_capped,
                   math = sat_math_4_capped)
```


```{r}
# visualize the distributions of scores
sat_df_4 %>% ggplot(aes(x = verbal)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Verbal Histogram",
       x = "Verbal Score",
       y = "Count")
```


```{r}
sat_df_4 %>% ggplot(aes(x = math)) +
  geom_histogram(color = "white") +
  labs(title = "Simulated SAT Math Histogram",
       x = "Math Score",
       y = "Count")

```

Alternatively, we could plot these together by either making a long data set or just overlaying histograms and modifying transparency (the alpha parameter)

```{r}
sat_df_4 %>% ggplot() +
  geom_histogram(aes(x = math, fill = "red"), color = "white", alpha = 0.2) +
  labs(title = "Simulated SAT Scores Histogram",
       x = "SAT Score",
       y = "Count") +
  geom_histogram(aes(x = verbal, fill = "blue"), color = "white", alpha = 0.2) +
  scale_fill_manual(name = "Category", values = c("blue", "red"), labels = c("verbal", "math"))

```

Try a boxplot instead

```{r}

sat_df_4 %>% 
  pivot_longer(cols = c("verbal", "math"), 
               names_to = "variable",
               values_to = "score") %>% 
  ggplot(aes(x = variable, y = score)) +
  geom_boxplot() +
  labs(x = "SAT Test Section",
       y = "Section Score",
       title = "Distribution of SAT Math and Verbal Scores")

```






Looking at the histogram, we can see that there appears to be a systematic difference in the scores. Running a t-test will help us see if that initial observation is statistically correct. 
Since we will be comparing verbal and math scores for each student, this will be a dependent samples (or paired samples) t-test

```{r}
sat_t_test_4 <- t.test(sat_df_4$math, sat_df_4$verbal, paired = TRUE) # don't forget to add: paired = TRUE
# now look at the results
sat_t_test_4
```



We should see that the t-test will pick up differences between means if there is a sufficiently large difference.



Calculate an effect size. Remember that $r = \sqrt{\frac{t^2}{t^2 + df}}$

```{r}

(t <- sat_t_test_4$statistic[[1]]) # wrapping this in parentheses prints out the value of t at the same time that it stores the value of t

(df <- sat_t_test_4$parameter[[1]])

(r <- sqrt(t^2/(t^2 + df)))
round(r,3)
r^2
```
