---
title: "Week_6_exercise"
author: "Katz"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(car)
library(psych)
```


## For this exercise, you will be working with the student_happiness.csv file.

This file contains simulated data with several variables. The scenario is the following: 
You are interested in studying engineering student well being. One dimension of student health is their mental health. In particular, you have a measure of student happiness. This will be your outcome variable.

You have reason to believe that there are several factors that can contribute to a person's happiness. In particular, you believe that their class standing (e.g., undergrad, masters, phd), discipline (you only sampled mechanical, civil, and electrical engineering students), time spent outdoors, and time spent on zoom all play a role in a student's happiness. Sooo you go out and survey students, collecting each of these variables in addition to giving them a series of questions that let you calculate their happiness score. The composite score is what you now have in your student_happiness.csv file.

The objective here is to model the outcome as a function of the different predictors that you have. There will be a little less scaffolding than last week, but you can follow the class example in the Week_6_demo if you get stuck or can't think of what to do next...


First, it's probably a good idea to load in the data...
```{r}

```


Next, it's never a bad idea to start summarizing what you have. You can try visualizing things with geom_bar() or geom_point() (as appropriate) or make summary tables with things like group_by() and summarize() or count().

```{r}

```

Now, it's probably time to start creating your model (or models...?)

```{r}

```


Check your residuals for outliers and other model diagnostics

```{r}

```



