---
title: "Week 3 Notes"
author: "Katz"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```



## One continuous variable (either predictor or outcome variable)

```{r example-hist, fig.show='hide', echo = TRUE}
num <- 50
mu <- 5
stdev <- 2

x <- rnorm(n = num, mean = mu, sd = stdev)
hist(x)


```


You can also do this using ggplot rather than base R graphics

```{r ggplot-hist}
num <- 50
mu <- 5
stdev <- 2

x_vec <- rnorm(n = num, mean = mu, sd = stdev)
x_df <- tibble(x_col = x_vec)

ggplot(data = x_df, mapping = aes(x = x)) +
  geom_histogram()

```


You can use geom_density() instead of geom_histogram() to get a smooth graph


```{r echo = FALSE}

num <- 50
mu <- 5
stdev <- 2

x_vec <- rnorm(n = num, mean = mu, sd = stdev)
x_df <- tibble(x_col = x_vec)



ggplot(data = x_df, mapping = aes(x = x_col)) +
  geom_density()


```



Just for fun, look at what happens to the the plot if you increase the sample size


```{r echo = FALSE}

num <- 5000
mu <- 5
stdev <- 2

x_vec <- rnorm(n = num, mean = mu, sd = stdev)
x_df <- tibble(x_col = x_vec)



ggplot(data = x_df, mapping = aes(x = x_col)) +
  geom_density()


```



## One Discrete Variable (either predictor or outcome)


```{r geom-bar}


```




## Discrete Predictor, Continuous Outcome


```{r boxplot-example}
group_size <- 20
chem_e_scores <- rnorm(n = group_size, mean = 85, sd = 4)
chem_scores <- rnorm(n = group_size, mean = 78, sd = 6)


data_df <- tibble(
  discipline = rep(c("ChemE", "Chemistry"), each = group_size),
  score = c(chem_e_scores, chem_scores)
)

data_df %>% 
  ggplot(aes(x = discipline, y = score)) +
  geom_boxplot()

```


You can make a few modifications to possibly make this easier to read. 

The first is to put the discrete category on the y axis instead of the x axis.

The second is to use geom_jitter() in addition to geom_boxplot() to show the individual points in each group. 


```{r box-and-dot-plot}

data_df %>% 
  ggplot(aes(x = score, y = discipline)) +
  geom_boxplot() +
  geom_jitter()

```




