---
title: "Week 5 Exercises"
author: "Katz"
date: "2/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are two exercises for practicing regression in this rmd file. The first involves using two data sets provided to you. The second exercise involves generating your own data as part of the process of gaining some intuition about what kind of underlying data generating process (there's that term from week 2!) might be creating the data you actually observe in your sample.

## Exercise 1 - Teacher salary linear regression demo

First, read in the data. In this case, the two CSV files are stored in a folder called "data". Be sure to adjust this path however you need.

```{r}
file_path_prin <- "./data/principalSalaries.csv"
file_path_tea <- "./data/teacherSalaries.csv"

principal_salaries <- read_csv(file_path_prin)
teacher_salaries <- read_csv(file_path_tea)
```

#### Try to filter out the public schools and just pick out the average salaries in the 2014-2016 year range (hint: this is stored as "Av14_16P" for principals.) and div_num column.

```{r}
## Your code here to filter and select the two columns
```

#### Convert the div_num column to numeric

```{r}
## Your code here

```

#### Repeat the same two steps above for teacher salaries (filter public schools, pick out the 2014-2016 salaries (using select()), and convert the div_num column to a numeric data type (using something like as.numeric())

```{r}
## Your code here for filtering, selecting, and converting data type 
```

#### Join the teacher salaries and principal salaries into one dataframe by using an inner_join. You should think about which column to join on (i.e., which to pass to the by = "" argument within inner_join())

```{r}
combined_salaries <- inner_join(###,
                                ###,
                                by=###)

```

This section helps make tables for formatting in r or printing to csv file to load into excel etc

```{r}
making_a_table <- describe(combined_salaries$Av14_16P) %>%
  select(mean,sd,skew,kurtosis)
```

```{r}
kable(making_a_table) %>% 
  kable_styling("striped", full_width = F) 
```

let's talk about running a linear model (simple regression) with the teacher

#### Create a linear model using the lm() function that models principal salary as a function of teacher salary (i.e., principal salary is the outcome variable). Replace the hashtags below

```{r}
fit1 <- lm(### ~ ###, data= ###)

summary(fit1)
```

Plot the principal salaries against the teacher salaries. Think about what kind of plot makes the most sense.

```{r}
### Your code here

```

## Exercise 2 - Create Your Own Adventure

For this exercise, first generate your own data by creating a series of outcome variable values and a series of predictor variable values. For each outcome you should have a predictor (and vice versa). That's just a long way of saying if you want to simulate having 50 observations then you should have 50 outcomes and 50 predictors.
